---
title: "Time Series Forecasting using VAR model"
author: "Swetha Ramanadham"
date: "1/14/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("vars")#install the package vars
library(vars)
```

#### Get a basic understanding of the dataset
```{r cars}
data(Canada)
dim(Canada)
colnames(Canada)
```

#### To know the arguments of VAR function
```{r pressure, echo=FALSE}
args(VAR)
```

```{r}
args(VARselect) #to know the arguments of VAR select function
```

#### Determining the best lags among 5 using varselect function
```{r}
VARselect(Canada, lag.max = 5, type = "const")
```
we have few ‘lag selection criteria’ like HQ, SIC, AIC and LR, etc.
Generally, we choose the lag length for which the values of most of these lag length criteria are minimized. According to the more conservative SC(n) and HQ(n) criteria, the empirical optimal lag-order is 2.




#### Building the VAR model using VAR() by specifying the no of lags
```{r}
var.2c <- VAR(Canada, p = 2, type = "const")
```


#### Testing causality to know if the real wage and productivity is causal to employment and unemployment.
```{r}
causality(var.2c, cause = c("rw", "prod"))
causality(var.2c, cause = c("prod", "rw"))
```
As in all the tests for granger as 'p' value is less than 0.05 we reject the null hypothesis

```{r}
par(mar=c(1,1,1,1))#run only if plots are not displaying because of an error
```


```{r}
plot(var.2c)
```

The above graph is a  representation of fit and residual of variable ‘e’

#### Summary of estimation results
```{r}
summary(var.2c)
```
Form the above tables we can see that each variable is a linear function of the past values of itself and the past values of all the other variables


#### Forecasting
```{r}

var.f10 <- predict(var.2c, n.ahead = 2, ci = 0.95) 
names(var.f10)
```


```{r fig.height=10}
plot(var.f10)
```


